\documentclass[a4paper,11pt,twoside]{memoir}
\chapterstyle{plain}

\usepackage{TUINFDA}

\usepackage{url}
\usepackage{hyperref}					% links in pdf
\usepackage{graphicx}            			% Figures
\usepackage{interval}
\usepackage{verbatim}            			% Code-Environment
\usepackage[lined,linesnumbered,algochapter]{algorithm2e} % Algorithm-Environment
\newsubfloat{figure}
\newsubfloat{table}

\usepackage{booktabs}
\usepackage{pgf}					
\usepackage{tikz}					% tikz graphics
\usetikzlibrary{arrows,automata}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{colortbl}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{groupplots}
\usetikzlibrary{pgfplots.groupplots}
%\usetikzlibrary{external}
\usetikzlibrary{matrix}
\usetikzlibrary{spy}
%\tikzexternalize[prefix=tikz/]
%\tikzset{external/up to date check=md5}

\usepackage{ngerman}
\usepackage[ngerman]{babel}
\usepackage{bibgerm,cite}       % Deutsche Bezeichnungen, Automatisches Zusammenfassen von Literaturstellen
\usepackage[ngerman]{varioref}  % Querverweise
% to use the german charset include cp850 for MS-DOS, ansinew for Windows and latin1 for Linux.
% \usepackage[latin1]{inputenc}
\usepackage{mathtools}
\usepackage{array}
\usepackage{multirow}
\usepackage{siunitx}
\usepackage{commath}
\usepackage{hhline}
\usepackage[sectionbib]{natbib}
\usepackage{csvsimple}

\DeclareSIUnit\pixel{px}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\newcommand{\mvec}[1]{\mathbf{#1}}
\newcommand{\mvecx}[1]{\mathbf{#1}_x}
\newcommand{\mvecy}[1]{\mathbf{#1}_y}
\newcommand{\mvecz}[1]{\mathbf{#1}_z}
\newcommand{\mvecw}[1]{\mathbf{#1}_w}
\newcommand{\mmat}[1]{\mathbf{#1}}
\newcommand{\transpose}[1]{#1^{\mathsf{T}}}
\newcommand{\inverse}[1]{#1^{\mathsf{-1}}}
\newcommand{\normalise}[1]{\frac{#1}{\norm{#1}}}

\newcommand{\wavevector}{wavevector\xspace}
\newcommand{\wavevectors}{wavevectors\xspace}
\newcommand{\Wavevector}{Wavevector\xspace}
\newcommand{\Wavevectors}{Wavevectors\xspace}
\newcommand{\wavenumber}{wavenumber\xspace}
\newcommand{\wavenumbers}{wavenumbers\xspace}
\newcommand{\Wavenumber}{Wavenumber\xspace}
\newcommand{\Wavenumbers}{Wavenumbers\xspace}
\newcommand{\wavelength}{wavelength\xspace}
\newcommand{\wavelengths}{wavelengths\xspace}
\newcommand{\Wavelength}{Wavelength\xspace}
\newcommand{\Wavelengths}{Wavelengths\xspace}

\newcommand{\FourierTransform}{Fourier Transform\xspace}
\newcommand{\FourierTransforms}{Fourier Transforms\xspace}
\newcommand{\InvFourierTransform}{Inverse Fourier Transform\xspace}
\newcommand{\InvFourierTransforms}{Inverse Fourier Transforms\xspace}
\newcommand{\DiscreteFourierTransform}{Discrete Fourier Transform\xspace}
\newcommand{\DiscreteFourierTransforms}{Discrete Fourier Transforms\xspace}
\newcommand{\InvDiscreteFourierTransform}{Inverse Discrete Fourier Transform\xspace}
\newcommand{\InvDiscreteFourierTransforms}{Inverse Discrete Fourier Transforms\xspace}
\newcommand{\FastFourierTransform}{Fast Fourier Transform\xspace}
\newcommand{\FastFourierTransforms}{Fast Fourier Transforms\xspace}
\newcommand{\FT}{FT\xspace}
\newcommand{\IFT}{IFT\xspace}
\newcommand{\DFT}{DFT\xspace}
\newcommand{\IDFT}{IDFT\xspace}
\newcommand{\IDFTs}{IDFTs\xspace}
\newcommand{\FFT}{FFT\xspace}
\hypersetup{draft}

\thesistitle{Ocean Surface Generation and Rendering}
%\thesissubtitle{Optional Subtitle} % optional
\thesisdate{11.07.2018}

% all titles and designations have to be gender-related!
\thesisdegree{Diplom-Ingenieur}{Diplom-Ingenieur}
\thesiscurriculum{Visual Computing}{Visual Computing} % your study
\thesisverfassung{Verfasser} % Verfasser
\thesisauthor{Thomas Gamper} % your name
\thesisauthoraddress{Lehmanngasse 25/2/12, 1230 Wien} % your address
\thesismatrikelno{0107543} % your registration number

\thesisbetreins{Associate Prof. Dipl.-Ing. Dipl.-Ing. Dr.techn. Michael Wimmer}
\thesisbetrzwei{}
%\thesisbetrdrei{Dr. Vorname Familienname} % optional

% define page numbering styles
\makepagestyle{numberCorner}
\makeevenfoot{numberCorner}{\thepage}{}{}
\makeoddfoot{numberCorner}{}{}{\thepage}

% define custom macros for specific formats or names
\newcommand{\uml}[1]{\texttt{#1}}
\newcommand{\cd}{\textsf{Class Diagram}}

\DeclareMathOperator{\sech}{sech}

\intervalconfig{
soft open fences,
}

\renewcommand{\thesection}{\arabic{section}}
\makeatletter
\let\l@section\l@chapter
\makeatother

\begin{document}
\captionnamefont{\bfseries}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   FRONTMATTER    %%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\frontmatter
\pagenumbering{roman}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   TITLEPAGES    %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% the german title page is required as first page
\include{titlepage}

% an english translation may follow
%\include{titlepage_en} % optional

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   ERKLAERUNG DER SELBSTAENDIGKEIT   %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\cleardoublepage
%\selectlanguage{ngerman}
%\include{chapters/erklaerung}
%\selectlanguage{english}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   ACKNOWLEDGEMENTS    %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% optional acknowledgements may be included in german or in english
%\input{chapters/danksagung} 		% optional
%\include{chapters/acknowledgements}	% optional

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   ABSTARCT    %%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\include{chapters/abstract_en}
%\cleardoublepage
%\selectlanguage{ngerman}
%\include{chapters/abstract_de}
%\selectlanguage{english}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   CONTENTS    %%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% uncomment to set document language to german (results in "Inhaltsverzeichnis", "Kapitel", "Abbildung", etc. instead of "Contents", "Chapter", and "Figure"), otherwise the document's language is english
%\selectlanguage{ngerman}

\setcounter{tocdepth}{1}

%\cleardoublepage
%\pagestyle{numberCorner}
%\tableofcontents*

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   MAINMATTER    %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\mainmatter
\pagenumbering{arabic}
\pagestyle{numberCorner}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\include{chapters/introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\include{chapters/state_of_the_art}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\include{chapters/background}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\include{chapters/implementation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\include{chapters/results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\include{chapters/summary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Structure}
%\label{ch:structure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\input{chapters/structure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Typographic Design}
%\label{ch:typo}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\input{chapters/design}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Bibliographic Issues}
%\label{ch:bibliographic}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\input{chapters/bibliographic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BACKMATTER %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\appendix
%\include{chapters/appendix}

\section{Problem Statement}
Natural phenomena are a challenging topic in the field of computer graphics.
Actual examples are complex structures which are to be found throughout nature,
such as mountains, trees and water. The latter is especially interesting
because of its highly dynamic form, which poses a variety of sophisticated problems.
For computer graphics to reproduce the diverse appearance of ocean surfaces
represents one such problem.
Rendering the ocean is a demanding task for several reasons. First, consider the
sheer size of a water body as large as the ocean, which in numerous viewing
situations will be visible all the way to the horizon. Second, the ocean surface
is dynamic, therefore it needs to be constantly updated with the passage of time
even though the wave interactions that define its shape are huge in terms of complexity.
% thus we employ approximative models. %published by the oceanographic research community.
Third, the optics of water are intricate. Incoming light may be reflected at the
surface or may be refracted into the water body, with some of the refracted light
possibly finding its way back to the ocean's surface, either by scattering inside the water
body or by reflection at the sea bottom. Moreover, waves on the ocean surface may
break and cause surf and foam, both of which strongly deviate in appearance from
the surrounding water surface.
But it would not be viable for one work to try to cover all the dynamics, as well
as the lighting, of the entire ocean. Therefore we may approach a more specific
subject, namely the synthesis of animated ocean surfaces which are both believable
and computationally feasible. 
Fortunately, the oceanographic research community already developed models
\citep[e.g.\ ][]{article:PiersonMoskowitz1964, article:Hasselman1973} which satisfy the
better part of those requirements, as they are essential for oceanographic simulations.
Still, to integrate said models into a coherent framework for ocean surface
synthesis may pose a significant challenge.
Moreover, only a representation of the oceanographic models most adequate for
computer graphics algorithms would allow us to reduce computational complexity to
a level where real-time operation may still be possible.
\section{Aim of the Work}
The aim of this work is to produce an optimized implementation which is able to generate,
animate and render the surface of an open ocean in real time. We focus our interest
on the synthesis of animated ocean surface geometry, for which we will adopt
a set of distinct models from oceanographic research. Said models may allow for
easy addition and reduction of detail, as well as for a range of algorithmic
optimizations. Thus, we may be able to improve upon the balance between model detail
and computational workload of current implementations.
%
%Whereas for ocean lighting
%purposes we seek to implement the approach by \citet{article:oceanlighting}, as it
%has specifically been tailored to the open ocean.
\section{Methodology}
The methodological approach consists of the following steps:
\begin{itemize}
\item Investigate existing ocean simulation and rendering methods.
\item Identify wave spectrum models from oceanographic research suitable for
ocean surface synthesis.
\item Research the theoretical background necessary to integrate said models
into a framework which gives coherent and consistent results.
\item Implement a real-time capable variant of the ocean surface synthesis algorithm by \citet{course:simulatingocean}.
\item Adopt the lighting algorithm by \citet{article:oceanlighting} for real-time ocean surface rendering with wave spectra.
\item Evaluate the obtained results with respect to the current state-of-the-art.
\end{itemize}
\section{State-of-the-Art}
%To date, computer graphics employs several ocean wave simulation models which
%may be roughly separated into three families: parametric description
%\citep{Biesel:1952}, spectral description \citep{book:kinsman2002wind}, and computational fluid dynamics \citep{Bridson:2015}. For our work we will focus on the spectral description,
%which approximates the ocean surface using wave spectra which simulate the sea
%as a random process based on the distribution of wave energy in frequency space.
%Given the wave spectrum, one is then able to synthesize the ocean surface
%by means of an \InvFourierTransform.
The seminal work by \citet{course:simulatingocean}
is a prominent example of wave spectrum based ocean simulation in computer
graphics. Therein, \citeauthor{course:simulatingocean}
describes how the ocean was brought to life for productions such as
\emph{Waterworld} and \emph{Titanic}.
The computer graphics community has been, and still is, hard at work to make
the algorithms presented by \citeauthor{course:simulatingocean} capable to render in real time.
The challenge is trifold: first, the synthesis of an animated ocean surface,
second, efficient rendering of vast ocean surface geometry, and third,
believable shading of the ocean surface.
Note that the spectral description of the ocean is an approximation
which simulates the sea as a random process based on the distribution of wave energy in frequency space. Given the wave spectrum, one is then able to synthesize the ocean surface by means of an \InvFourierTransform. The latter exists
as a highly optimized algorithm \citep{Cooley:1965} which, depending on
input resolution, may allow for real-time operation on nowadays hardware
(CPU and GPU). With regard to the vast ocean geometry, \citet{thesis:johanson} presents a constant-overhead, adaptive meshing scheme which has been tailored to large, animated water surfaces.
Last, with regard to believable shading of the water surface, \citet{article:oceanlighting} describe a real-time rendering algorithm
for seamless ocean surface lighting. Thus, we should be able to reach our
goal by combining these existing techniques in a sensible manner.
\section{Relation to Visual Computing}
The curriculum for \emph{Visual Computing} touches the topics of natural phenomena,
water rendering, and real-time rendering in several courses. The most directly
linked courses are the following:
\begin{itemize}
\item 186.101 Rendering
\item 186.140 Real-Time Rendering
\item 186.192 Algorithms for Real-Time Rendering
\item 186.830 Computergraphics
\item 186.831 Computergraphics 
\item 186.120 Fractals
\item 186.186 Fractals
\end{itemize}

%\backmatter
%\listoffigures
%\listoftables
%\listofalgorithms

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
